### 2023-02-05 talk/references.tex
add references I will need

% inputed at end of slides.tex

\newcommand{\shref}[2]{\,{\tiny \href{#1}{#2}}}
\begin{itemize}
\item[] A.~Brandt (1977). \emph{Multi-level adaptive solutions to boundary-value problems}, Mathematics of Computation 31 (138), 333--390
\item[] A.~Brandt \& C.~Cryer (1983). \emph{Multigrid algorithms for the solution of linear complementarity problems \dots}, SIAM J.~Sci.~Stat.~Comput.~4 (4), 655--684 \shref{https://doi.org/10.1137/0904046}{doi:10.1137/0904046}
%FTITLE Multigrid algorithms for the solution of linear complementarity problems arising from free boundary problems
\item[] E.~Bueler (2021). \emph{Conservation laws for free-boundary fluid layers}, SIAM J.~Appl.~Math.~81 (5), 2007--2032 \shref{https://doi.org/10.1137/20M135217X}{doi:10.1137/20M135217X}
\item[] E.~Bueler (2021). \emph{PETSc for Partial Differential Equations}, SIAM Press, Philadelphia
%\item[] E.~Bueler (2022). \emph{Performance analysis of high-resolution ice-sheet simulations}, J.~Glaciol., \href{https://doi.org/10.1017/jog.2022.113}{doi:10.1017/jog.2022.113}
\item[] P.~Farrell, M.~Croci, \& T.~Surowiec (2020). \emph{Deflation for semismooth equations}, Optimization Methods \& Software 35 (6), 1248--1271 \shref{https://doi.org/10.1080/10556788.2019.1613655}{doi:10.1080/10556788.2019.1613655}
\item[] C.~Gr{\"a}ser \& R.~Kornhuber (2009). \emph{Multigrid methods for obstacle problems}, J.~Comput.~Math., 1--44
\item[] G. Jouvet \& E. Bueler (2012). \emph{Steady, shallow ice sheets as obstacle problems \dots}, SIAM J.~Appl.~Math.~72 (4), 1292--1314 \shref{https://doi.org/10.1137/110856654}{doi:10.1137/110856654}
%FTITLE Steady, shallow ice sheets as obstacle problems: well-posedness and finite element approximation
\item[] D. Kinderlehrer \& G. Stampacchia (1980). \emph{An Introduction to Variational Inequalities and their Applications}, Academic Press, New York
%\item[] T.~Isaac, G.~Stadler, \& O.~Ghattas (2015). \emph{Solution of nonlinear Stokes equations \dots, with application to ice sheet dynamics}, SIAM J.~Sci.~Comput.~37 (6), B804--B833 \shref{https://doi.org/10.1137/140974407}{doi:10.1137/140974407}
%FTITLE Solution of nonlinear Stokes equations discretized by high-order finite elements on nonconforming and anisotropic meshes, with application to ice sheet dynamics
\item[] U.~Trottenberg, C.~Oosterlee, \& A. Schuller (2001).  \emph{Multigrid}, Elsevier, Oxford
\end{itemize}