### 2023-02-06 talk/background.tex
progress talk, plus reference structure

% inputed at end of slides.tex

\newcommand{\sdoi}[1]{\,{\tiny \href{https://doi.org/#1}{doi:#1}}}
\begin{itemize}
\item[] A.~Brandt (1977). \emph{Multi-level adaptive solutions to boundary-value problems}, Mathematics of Computation 31 (138), 333--390
\item[] E.~Bueler (2021). \emph{PETSc for Partial Differential Equations}, SIAM Press, Philadelphia
\item[] N.~Kikuchi \& J.~Oden (1988).  \emph{Contact Problems in Elasticity: A Study of Variational Inequalities and Finite Element Methods}, SIAM Press, Philadelphia
\item[] D.~Kinderlehrer \& G.~Stampacchia (1980). \emph{An Introduction to Variational Inequalities and their Applications}, Academic Press, New York
\item[] U.~Trottenberg, C.~Oosterlee, \& A. Schuller (2001).  \emph{Multigrid}, Elsevier, Oxford
\end{itemize}