### 2023-02-07 talk/slides.tex
progress ... now dinner

\documentclass[svgnames,
               hyperref={colorlinks,citecolor=DeepPink4,linkcolor=FireBrick,urlcolor=Maroon},
               usepdftitle=false]  % see \hypersetup{} below
               {beamer}

\mode<presentation>{
  \usetheme{Madrid}
  %\usecolortheme{seagull}
  \usecolortheme{seagull}
  \setbeamercovered{transparent}
  \setbeamerfont{frametitle}{size=\large}
}

\setbeamercolor*{block title}{bg=red!10}
\setbeamercolor*{block body}{bg=red!5}

%\usepackage[svgnames]{xcolor}
\usepackage{hyperref}
\hypersetup{
    pdftitle = {Toward nonlinear multigrid for nonlinear variational inequalities},
    pdfauthor = {Ed Bueler},
    pdfsubject = {},
    pdfkeywords = {}
}

\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{times}
\usepackage[T1]{fontenc}
\usepackage{empheq,bm,xspace,fancyvrb,soul}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows.meta,decorations.markings,decorations.pathreplacing,fadings,positioning}
\usepackage[kw]{pseudo}
\pseudoset{left-margin=15mm,topsep=5mm,label=,idfont=\texttt,st-left=,st-right=}

\makeatletter
%\newcommand\notsotiny{\@setfontsize\notsotiny\@vipt\@viipt}
\newcommand\notsotiny{\@setfontsize\notsotiny\@viipt\@viiipt}
\makeatother

\newcommand{\eps}{\epsilon}
\newcommand{\RR}{\mathbb{R}}

\newcommand{\grad}{\nabla}
\newcommand{\Div}{\nabla\cdot}
\newcommand{\trace}{\operatorname{tr}}

\newcommand{\hbn}{\hat{\mathbf{n}}}

\newcommand{\bb}{\mathbf{b}}
\newcommand{\be}{\mathbf{e}}
\newcommand{\bbf}{\mathbf{f}}
\newcommand{\bg}{\mathbf{g}}
\newcommand{\bn}{\mathbf{n}}
\newcommand{\br}{\mathbf{r}}
\newcommand{\bu}{\mathbf{u}}
\newcommand{\bv}{\mathbf{v}}
\newcommand{\bw}{\mathbf{w}}
\newcommand{\bx}{\mathbf{x}}

\newcommand{\bF}{\mathbf{F}}
\newcommand{\bQ}{\mathbf{Q}}
\newcommand{\bU}{\mathbf{U}}
\newcommand{\bV}{\mathbf{V}}
\newcommand{\bX}{\mathbf{X}}

\newcommand{\btau}{\bm{\tau}}
\newcommand{\bxi}{\bm{\xi}}

\newcommand{\bzero}{\bm{0}}

\newcommand{\rhoi}{\rho_{\text{i}}}

\newcommand{\ip}[2]{\left<#1,#2\right>}

\newcommand{\nn}{{\text{n}}}
\newcommand{\pp}{{\text{p}}}
\newcommand{\qq}{{\text{q}}}
\newcommand{\rr}{{\text{r}}}

\newcommand{\bus}{\bu|_s}
\newcommand{\oo}[1]{\displaystyle O\left(#1\right)}
\newcommand{\sold}{s_{\text{o}}}

\newcommand{\maxR}{R^{\bm{\oplus}}}
\newcommand{\minR}{R^{\bm{\ominus}}}
\newcommand{\iR}{R^{\bullet}}

\title[Multigrid for nonlinear VI]{Toward nonlinear multigrid \\ for nonlinear variational inequalities}

%\subtitle{\emph{x}}

\author{Ed Bueler}

\institute[UAF]{University of Alaska Fairbanks}

\date[]{February 2023}

%\titlegraphic{\begin{picture}(0,0)
%    \put(0,180){\makebox(0,0)[rt]{\includegraphics[width=4cm]{figs/software.png}}}
%  \end{picture}
%}

\titlegraphic{\hfill \includegraphics[width=0.15\textwidth]{images/uafbw.png}}

%% to start section counter at 0 see
%% https://tex.stackexchange.com/questions/170222/change-the-numbering-in-beamers-table-of-content

\begin{document}
\beamertemplatenavigationsymbolsempty

%\begin{frame}
%  \maketitle
%\end{frame}

{
  %\usebackgroundtemplate{\includegraphics[width=\paperwidth]{images/gray-british-clark2022.png}}
  \begin{frame}
    \titlepage
  \end{frame}
}

\begin{frame}{Outline}
  \tableofcontents[hideallsubsections]
\end{frame}

\section{variational inequalities (VIs)?}

\begin{frame}{example: a classical obstacle problem}

\includegraphics[width=0.55\textwidth]{images/obstacle65.pdf} \qquad \includegraphics[width=0.35\textwidth]{images/obstacle-sets.png}

\bigskip
\only<1>{
\begin{itemize}
\item \emph{problem.} on a domain $\Omega \subset \RR^2$, find the displacement $u(x)$ of a membrane, with fixed value $u = g$ on $\partial \Omega$, above an \emph{obstacle} $\psi(x)$, which minimizes the elastic energy
    $$J(v) = \int_\Omega \frac{1}{2} |\grad v|^2 - f\, v$$
\item shown above: \quad $\Omega=(-2,2)^2$, $\psi(x)$ a hemisphere, $f(x)=0$
\end{itemize}

\phantom{x}
}
\only<2>{
\begin{itemize}
\item i.e.~constrained optimization over an \emph{admissible set}
	$$\mathcal{K} = \left\{v \in H^1(\Omega) \,:\, v\big|_{\partial \Omega} = g \text{ and } v \ge \psi\right\}$$
\item $J'(u)$ points directly into $\mathcal{K}$, the {\color{FireBrick} \emph{variational inequality} (VI)}:
    $$\ip{J'(u)}{v-u} = \int_\Omega \grad u\cdot \grad (v-u) - f (v-u) \ge 0 \quad \text{for all } v \in \mathcal{K}$$
\end{itemize}
}
\only<3>{
\begin{itemize}
\item the solution defines \emph{active} $A_u = \{u = \psi\}$ and \emph{inactive} $R_u = \{u> \psi\}$ subsets of $\Omega$, and a \emph{free boundary} $\Gamma_u=\partial R_u \cap \Omega$
\item naive strong form would pose the problem in terms of its solution:
\begin{align*}
-\grad^2 u &= f \quad \text{ on $R_u$} \\
u &= \psi \quad \text{ on $A_u$}
\end{align*}

\phantom{x}

\phantom{x}
\end{itemize}
}
\only<4>{
\begin{itemize}
\item the \emph{complementarity problem} (CP) is meaningful as a strong form:
\begin{align*}
u - \psi &\ge 0 \\
-\grad^2 u - f &\ge 0 \\
(u - \psi)(-\grad^2 u - f) &= 0
\end{align*}
   \begin{itemize}
   \item[$\circ$] for optimization problems: \quad CP $=$ KKT conditions
   \end{itemize}

\phantom{x}

\phantom{x}
\end{itemize}
}
\end{frame}

\begin{frame}{general variational inequalities}

\begin{itemize}
\item let $\mathcal{K}$ be a closed and convex subset of a Banach space $\mathcal{V}$
\item suppose $F:\mathcal{K} \to \mathcal{V}'$ is a continuous, generally nonlinear operator
    \begin{itemize}
    \item[$\circ$] $F$ may be defined only on $\mathcal{K}$
    \item[$\circ$] $F$ may \emph{not}\, be the derivative of an objective function $J$
    \end{itemize}
\item the general problem {\color{FireBrick} VI($F$,$\mathcal{K}$)} is
	$${\color{FireBrick} \ip{F(u)}{v-u} \ge 0 \quad \text{ for all } v \in \mathcal{K}}$$
\end{itemize}
\end{frame}

\begin{frame}{VI $=$ ``constrained equation''}

\begin{center}
\begin{tabular}{l|l}
\begin{minipage}[t][16mm][t]{0.4\textwidth}
unconstrained optimization:
$$\min_{u\in\mathcal{V}} J(u)$$
\end{minipage}
&
\begin{minipage}[t][16mm][t]{0.4\textwidth}
constrained optimization:
$$\min_{u\in\mathcal{K}} J(u)$$
\end{minipage}
\\ \hline
\begin{minipage}[t][16mm][t]{0.4\textwidth}
equation for $u \in \mathcal{V}$: {\LARGE \strut}

$$F(u)=0$$
\end{minipage}
&
\begin{minipage}[t][16mm][t]{0.4\textwidth}

\vspace{-2mm}
{\color{FireBrick} VI} for $u \in \mathcal{K}$:
$${\color{FireBrick} \ip{F(u)}{v-u} \ge 0 \quad \forall v \in \mathcal{K}}$$
\end{minipage}
\end{tabular}
\end{center}
\end{frame}

\begin{frame}{applications of VIs}

\begin{itemize}
\item elastic contact, Signorini problems (e.g.~Kikuchi \& Oden 1988)
\item viscous contact problems (de Diego et al.~2022)
\item pricing of American options in the Black-Scholes model
\item the geometry of glaciers \hfill $\longleftarrow$ \emph{more soon}

\bigskip
\item first-semester calculus
\end{itemize}

\begin{center}
\includegraphics[width=0.9\textwidth]{images/calcone.png}
\end{center}
\end{frame}

\AtBeginSection[]
{
  \begin{frame}{Outline}
    \tableofcontents[currentsection]
  \end{frame}
}

\section{full approximation scheme (FAS) multigrid for PDEs?}

\begin{frame}{nonlinear 2-grid scheme}

\begin{center}
$\Omega^h$\, \includegraphics[height=0.16\textheight]{images/fine-grid.png} \hspace{25mm} \includegraphics[height=0.16\textheight]{images/coarse-grid.png} \,$\Omega^H$
\end{center}

\only<1>{
\begin{itemize}
\item consider a nonlinear elliptic PDE problem:
	$$F(u) = \ell$$

	\begin{itemize}
	\item for example, $F : \mathcal{V} \to \mathcal{V}'$ for $\mathcal{V}=H^1(\Omega)$, with $\ell\in \mathcal{V}'$
	\end{itemize}
\item FEM/FD discretization gives algebraic system on fine grid $\Omega^h$:
    $$F^h(u^h) = \ell^h$$
\item suppose $w^h$ yields residual norm $\|\ell^h - F^h(w^h)\| > \text{TOL}$
\end{itemize}

\phantom{x}
}
\only<2-3>{
\begin{itemize}
\item how can we improve $w^h$ \emph{without} globally linearizing $F^h$?

are there alternatives to Newton's method?
\item note the \emph{residual} $r^h(w^h) = \ell^h - F^h(w^h)$ is computable,

while the \emph{error} $e^h = w^h-u^h$ is unknown
\item the residual definition can be rewritten
    $$F^h(w^h) - F^h(u^h) \stackrel{\ast}{=} -r^h(w^h)$$
\only<2>{\item for $F^h$ linear, try to solve this \emph{error eqn} $F^h(e^h) = -r^h(w^h)$ for $\tilde e^h$, and correct $w^h \leftarrow w^h-\tilde e^h$ to improve $w^h$?}
\only<3>{\item \st{for $F^h$ linear, try to solve this \emph{error eqn} $F^h(e^h) = -r^h(w^h)$ for $\tilde e^h$, and correct $w^h \leftarrow w^h-\tilde e^h$ to improve $w^h$?}}
\end{itemize}
}
\only<4>{
\begin{itemize}
\item \emph{nodewise problem}: for $\psi_j^h$ a hat function or dof, solve for $c\in\RR$:
	$${\color{FireBrick} \phi_j(c) = r^h(w^h + c \psi_j^h)[\psi_j^h] = 0}$$
\item sweeping through and solving nodewise problems is a \emph{smoother}
    \begin{itemize}
    \item[$\circ$] Fourier analysis on linear PDEs shows smoothing property
    \item[$\circ$] post-smoothing, $w^h-u^h$ and $r^h(w^h)$ have small high-frequency
    \end{itemize}
\item post-smoothing, both sides of $\ast$ should be accurately approximate-able on a coarser grid
\end{itemize}
}
\only<5>{
\begin{itemize}
\item Brandt (1977) proposed to replace $F^h(w^h) - F^h(u^h) = -r^h(w^h)$ with a coarse-grid, re-discretized version
\item \emph{full approximation storage} (FAS) \emph{equation}:
	$${\color{FireBrick} F^H(w^H) - F^H(\iR w^h) = -R r^h(w^h)}$$

    \begin{itemize}
    \item[$\circ$] $\iR:\mathcal{V}^h \to \mathcal{V}^H$ is \emph{injection}
    \item[$\circ$] $R:(\mathcal{V}^h)' \to (\mathcal{V}^H)'$ is \emph{canonical restriction}
    \item[$\circ$] if $w^h=u^h$ exactly then $w^H = \iR w^h$
    \end{itemize}

FIXME CHECK CAREFULLY FROM HERE
\item equivalently ${\color{FireBrick} F^H(w^H) = \ell^H}$ where ${\color{FireBrick} \ell^H = FIXME R\,F^h(u^h) - \iR\, r^h(w^h)}$
\end{itemize}
}
\end{frame}

\begin{frame}{FAS 2-grid solver}

\begin{align*}
&\text{smooth by sweeps over $j$:}  & &w^h \leftarrow \left[\phi_j(c) = 0 \,\forall j\right] \\
&\text{evaluate:}                   & &\ell^H = R\,F^h(u^h) - \iR\, r^h(w^h) \\
&\text{solve:}                       & &F^H(w^H) = \ell^H \\
&\text{correct}:                    & &w^h \leftarrow w^h + P(w^H - \iR w^h) \\
&\text{smooth by sweeps over $j$:}  & &w^h \leftarrow \left[\phi_j(c) = 0 \,\forall j\right]
\end{align*}

\bigskip
\begin{itemize}
\item evaluate$+$solve$+$correct \, $=$ \, \emph{coarse grid correction}
\end{itemize}
\end{frame}

\begin{frame}{nonlinear multigrid = FAS V-cycle or F-cycle}

\includegraphics[height=0.15\textheight]{images/mg-grids.png}

\bigskip

\hfill \mbox{
\includegraphics[height=0.15\textheight]{images/mg-vcycle.png} \quad
\includegraphics[height=0.15\textheight]{images/mg-fcycle.png}
}

\vspace{-3mm}

\only<1>{
{\small
\begin{pseudo}
\pr{fas-vcycle}: \\+
    for $j=J$ downto $j=1$ \\+
      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j;u^j)$ \\
      $u^{j-1} = \iR u^j$ \\
      $\ell^{j-1} = f^{j-1}(u^{j-1}) + R \left(\ell^j - f^j(u^j)\right)$ \\-
    $\text{\pr{solve}}(\ell^0;u^0)$ \\
    for $j=1$ to $j=J$ \\+
      $u^j \gets u^j + P (u^{j-1} - \iR u^j)$ \\
      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j;u^j)$ \\-
\end{pseudo}
}
}
\only<2>{
\begin{itemize}
\item x
\end{itemize}
}
\end{frame}

\begin{frame}{nonlinear elliptic PDE problem}

\begin{itemize}
\item x
\end{itemize}
\end{frame}

\begin{frame}{Bratu model problem: optimality}

\begin{columns}
\begin{column}{0.6\textwidth}
\begin{itemize}
\item x
\end{itemize}
\end{column}
\begin{column}{0.4\textwidth}
\includegraphics[width=\textwidth]{images/bratu-exps.png}

\includegraphics[width=\textwidth]{images/bratu-time.png}
\end{column}
\end{columns}
\end{frame}

\section{the nonlocal VI for a fluid layer in a climate}

\begin{frame}{problem: fluid layer in a climate}

\begin{center}
\includegraphics[width=0.6\textwidth]{images/fluid-in-climate.png}
\end{center}

\begin{itemize}
\item x
\end{itemize}
\end{frame}

\section{a possibly scalable approach to nonlocal VIs}

\begin{frame}{Newton iterations for VIs}

\begin{columns}
\begin{column}{0.6\textwidth}
\begin{itemize}
\item x
\end{itemize}
\end{column}
\begin{column}{0.4\textwidth}
\includegraphics[width=\textwidth]{images/vi-newton-gmg-bad.png}

\includegraphics[width=\textwidth]{images/vi-newton-gmg-good.png}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{Newton-multigrid for VIs}

xxx
\end{frame}

\begin{frame}{optimality}

\begin{columns}
\begin{column}{0.6\textwidth}
\begin{itemize}
\item x
\end{itemize}
\end{column}
\begin{column}{0.4\textwidth}
\includegraphics[width=\textwidth]{images/obstacle-flops-per-n.png}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{the Stokes problem for ice}

\begin{itemize}
\item whether explicit or implicit, a non-shallow model solves a Stokes problem at each step:
\begin{align*}
- \nabla \cdot \left(2 \nu_\eps(D\bu)\, D\bu\right) + \nabla p - \rhoi \mathbf{g} &= \bzero && \text{in $\Lambda$} \\
\nabla \cdot \bu &= 0 && \text{''} \\
\btau_b - \bbf(\bu|_b) &= \bzero && \text{on $\Gamma_b$} \\
\bu|_b \cdot \bn_b &= 0 && \text{''} \\
\left(2 \nu_\eps(D\bu) D\bu - pI\right) \bn &= \bzero && \text{on $\Gamma_s$}
\end{align*}
\item this is the \alert{stress balance} (conservation of momentum) problem which determines velocity $\bu$ and pressure $p$
\item how fast is the numerical solution process?
    \begin{itemize}
    \item[$\circ$] how do solution algorithms \alert{scale with increasing resolution}?
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{multilevel NCP-coupled-to-Stokes solvers}

\begin{itemize}
\item direct attack on the problem seems to require a \alert{multilevel} solver for \alert{variational inequalities} (VIs), but in the \alert{non-local residual case}
    \begin{itemize}
    \item[$\circ$] this seems not to exist
    \item[$\circ$] the \alert{smoother} must compute an implicit-step SKE residual including surface-motion term $\Phi(s) = - \bu|_s\cdot \bn_s$ from a scalable Stokes solver
    \end{itemize}
\item near-optimal multilevel VI solvers exist for toy problems
    \begin{itemize}
    \item[$\circ$] Poisson equation obstacle problem (Brandt \& Cryer 1983; Gr\"aeser \& Kornhuber 2009)
    \end{itemize}
\end{itemize}
\end{frame}

\section{conclusion}

\begin{frame}{\alert{summary}}

\begin{itemize}
\item x
\end{itemize}
\end{frame}

\begin{frame}{references}

{\scriptsize
%{\notsotiny
\input{references.tex}
}
\end{frame}

\begin{frame}{\emph{background} references}

{\scriptsize
%{\notsotiny
\input{background.tex}
}
\end{frame}

\end{document}