### 2023-02-08 talk/references.tex
some progrees

% inputed at end of slides.tex

\newcommand{\sdoi}[1]{\,{\tiny \href{https://doi.org/#1}{doi:#1}}}
\begin{itemize}
\item[] A.~Brandt \& C.~Cryer (1983). \emph{Multigrid algorithms for the solution of linear complementarity problems \dots}, SIAM J.~Sci.~Stat.~Comput.~4 (4), 655--684 \sdoi{10.1137/0904046}
%FTITLE Multigrid algorithms for the solution of linear complementarity problems arising from free boundary problems
\item[] P.~Brune, M.~Knepley, B.~Smith, \& X.~Tu (2015). \emph{Composing scalable nonlinear algebraic solvers}, SIAM Review~57 (4), 535--565 \sdoi{10.1137/130936725}
\item[] E.~Bueler (2021). \emph{Conservation laws for free-boundary fluid layers}, SIAM J.~Appl.~Math.~81 (5), 2007--2032 \sdoi{10.1137/20M135217X}
%\item[] E.~Bueler (2022). \emph{Performance analysis of high-resolution ice-sheet simulations}, J.~Glaciol., \sdoi{10.1017/jog.2022.113}
\item[] G.~de Diego, P.~Farrell, \& I.~Hewitt (2022). \emph{Numerical approximation of viscous contact problems applied to glacial sliding}, J.~Fluid Mech.~938 (A21). \sdoi{10.1017/jfm.2022.178}
\item[] P.~Farrell, M.~Croci, \& T.~Surowiec (2020). \emph{Deflation for semismooth equations}, Optimization Methods \& Software 35 (6), 1248--1271 \sdoi{10.1080/10556788.2019.1613655}
\item[] C.~Gr{\"a}ser \& R.~Kornhuber (2009). \emph{Multigrid methods for obstacle problems}, J.~Comput.~Math., 1--44
\item[] T.~Isaac, G.~Stadler, \& O.~Ghattas (2015). \emph{Solution of nonlinear Stokes equations \dots ice sheet dynamics}, SIAM J.~Sci.~Comput., 37 (6), B804--B833 \sdoi{10.1137/140974407}
%FTITLE Solution of nonlinear Stokes equations discretized by high-order finite elements on nonconforming and anisotropic meshes, with application to ice sheet dynamics

\item[] G. Jouvet \& E. Bueler (2012). \emph{Steady, shallow ice sheets as obstacle problems \dots}, SIAM J.~Appl.~Math.~72 (4), 1292--1314 \sdoi{10.1137/110856654}
%FTITLE Steady, shallow ice sheets as obstacle problems: well-posedness and finite element approximation
\item[] A.~Reusken (1987). \emph{Convergence of the multigrid full approximation scheme \dots}, Numer.~Math.~52, 251--277 \sdoi{10.1007/BF01398879}
%FTITLE Convergence of the multigrid full approximation scheme for a class of elliptic mildly nonlinear boundary value problems
\item[] J.~Seguinot \& 5 others (2018).  \emph{Modelling last glacial cycle ice dynamics in the Alps}, The Cryosphere, 12 (10), 3265--3285 \sdoi{10.5194/tc-12-3265-2018}
%FAUTHORS Seguinot, J., Ivy-Ochs, S., Jouvet, G., Huss, M., Funk, M., & Preusser, F.
\end{itemize}