### 2023-06-12 paper/tikz/imagesvcycle.tex
visualize a V-cycle; see comments before fig:imagesvcycle regarding generation

\begin{tikzpicture}[scale=1.0]
    \pgfmathsetmacro{\HEIGHT}{2}
    \pgfmathsetmacro{\XSTEP}{3.7}
    \pgfmathsetmacro{\YSTEP}{3.7}
    \pgfmathsetmacro{\YLABEL}{-0.4}
    \pgfmathsetmacro{\NUDGE}{0.5}
    \node (start3) [] at (0,3*\YSTEP) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/start3.png}};
    \node (start3label) [below of = start3, yshift=\YLABEL cm] {$w^3 \ge \underline{\gamma}^3$};
    \node (down3) [] at (1.0*\XSTEP,3*\YSTEP) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/down3.png}};
    \node (down3label) [below of = down3, yshift=\YLABEL cm] {$y^3 \ge \underline{\phi}^3$};
    \node (down2) [] at (1.2*\XSTEP,2*\YSTEP) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/down2.png}};
    \node (down2label) [below of = down2, yshift=\YLABEL cm] {$y^2 \ge \underline{\phi}^2$};
    \node (down1) [] at (1.4*\XSTEP,1*\YSTEP) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/down1.png}};
    \node (down1label) [below of = down1, yshift=\YLABEL cm] {$y^1 \ge \underline{\phi}^1$};
    \node (up0) [] at (1.9*\XSTEP,0) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/up0.png}};
    \node (up0label) [below of = up0, yshift=\YLABEL cm] {$z^0 \ge \underline{\chi}^0$};
    \node (up1) [] at (2.4*\XSTEP,1*\YSTEP) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/up1.png}};
    \node (up1label) [below of = up1, yshift=\YLABEL cm] {$z^1 \ge \underline{\chi}^1$};
    \node (up2) [] at (2.6*\XSTEP,2*\YSTEP) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/up2.png}};
    \node (up2label) [below of = up2, yshift=\YLABEL cm] {$z^2 \ge \underline{\chi}^2$};
    \node (up3) [] at (2.8*\XSTEP,3*\YSTEP) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/up3.png}};
    \node (up3label) [below of = up3, yshift=\YLABEL cm] {$z^3 \ge \underline{\chi}^3$};
    \node (next3) [] at (3.8*\XSTEP,3*\YSTEP) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/next3.png}};
    \node (next3label) [below of = next3, yshift=\YLABEL cm] {$\tilde{w}^3 \ge \underline{\gamma}^3$};
    \graph [edges = {very thick}] {
        (start3) -> (down3)
                 -!- [] (down3label) -> (down2)
                 -!- [] (down2label) -> (down1)
                 -!- [] (down1label) -> (up0)
                 -> (up1label) -!- (up1)
                 -> (up2label) -!- (up2)
                 -> (up3label) -!- (up3)
                 -> (next3)
    };
\end{tikzpicture}