### 2023-07-14 paper/tikz/directRbadcoarse.tex
do a PEF (color!) and clean some text

\begin{tikzpicture}[scale=1.5]
  \draw[blue,thick] (0.0,1.5) -- (4.0,1.5)
                     node [xshift=-6mm,yshift=4mm] {{\color{black} $\underline{\gamma}^0 = \maxR\underline{\gamma}^1$?}};
  \draw[red,thick] (0.0,1.0) -- (4.0,1.0)
                     node [xshift=-6mm,yshift=-4mm] {{\color{black} $\overline{\gamma}^1 = \minR\overline{\gamma}^1$?}};
  \draw[black,thin,->] (-0.3,0.0) -- (4.3,0.0) node [xshift=5mm] {$x$};
  \foreach \x in {0.0, 2.0, 4.0} {
      \filldraw (\x,0.0) circle (1.5pt);
  }
\end{tikzpicture}