### 2023-07-17 paper/tikz/directRbadcoarse.tex
Another editing pass

\definecolor{seabornblue}{rgb}{0.2980392156862745, 0.4470588235294118, 0.6901960784313725}
\definecolor{seaborngreen}{rgb}{0.3333333333333333, 0.6588235294117647, 0.40784313725490196}
\definecolor{seabornred}{rgb}{0.7686274509803922, 0.3058823529411765, 0.3215686274509804}

\begin{tikzpicture}[scale=1.5]
  \draw[seabornblue,very thick] (0.0,1.5) -- (4.0,1.5)
                     node [xshift=-6mm,yshift=4mm] {{\color{black} $\underline{\gamma}^0 = \maxR\underline{\gamma}^1$?}};
  \draw[seabornred,very thick] (0.0,1.0) -- (4.0,1.0)
                     node [xshift=-6mm,yshift=-4mm] {{\color{black} $\overline{\gamma}^1 = \minR\overline{\gamma}^1$?}};
  \draw[black,thin,->] (-0.3,0.0) -- (4.3,0.0) node [xshift=5mm] {$x$};
  \foreach \x in {0.0, 2.0, 4.0} {
      \filldraw (\x,0.0) circle (1.5pt);
  }
\end{tikzpicture}