### 2024-07-09 talk-ismp/references.tex
start on slides for Montreal

% inputed at end of slides.tex

\newcommand{\sdoi}[1]{\,{\tiny \href{https://doi.org/#1}{doi:#1}}}
\begin{itemize}
\item E.~Bueler \& P.~Farrell (2024).  \emph{A full approximation scheme multilevel method for nonlinear variational inequalities}, to appear SIAM J.~Sci.~Comput. \, {\tiny \href{https://arxiv.org/abs/2308.06888}{arxiv:2308.06888}}
\item C.~Gr{\"a}ser \& R.~Kornhuber (2009). \emph{Multigrid methods for obstacle problems}, J.~Comput.~Math., 1--44
\item G.~Jouvet \& E.~Bueler (2012). \emph{Steady, shallow ice sheets as obstacle problems: well-posedness and finite element approximation}, SIAM J.~Appl.~Math.~72 (4), 1292--1314 \sdoi{10.1137/110856654}
\item X.~Tai (2003). \emph{Rate of convergence for some constraint decomposition methods for nonlinear variational inequalities}, Numer.~Math.~93 (4), 755--786 \sdoi{10.1007/s002110200404}
\end{itemize}